An Experimental Evaluation of Hybrid Vector Search (EA\&B)

### 2025-04-28 VLDB2026/main.tex
modify

	\noindent\textbf{\underline{UNG \cite{UNG}.}} The above simultaneous filtering algorithms may not guarantee the completeness of the results and perform poorly when the attribute selectivity is low. UNG is proposed to overcome these limitations.
	UNG is a unified framework that integrates diverse graph-based ANN indexes for hybrid query. It first groups the dataset by attribute sets, ensuring that vectors in each group share identical attributes. Then, it constructs a Label Navigating Graph (LNG) to encode inclusion relationships among attribute sets. Within each group, UNG builds graph-based ANN indexes (e.g., Vamana, HNSW) and connects them via cross-group edges to enable efficient cross-group search.
	UNG supports Boolean filtering with 2 modes:  1) The query attribute set is a subset of the data attribute set.  2) The query attribute set exactly matches the data attribute set.

@@

@@

	For attribute filtering tasks, we employ 7 real-world datasets widely adopted in existing literature: Msong~\cite{msong2011}, Audio~\cite{audio_unknown}, SIFT1M~\cite{sift2010}, GIST1M~\cite{sift2010}, GloVe~\cite{GloVe2015}, Enron~\cite{enron2015} and Text2Image~\cite{texttoimage}, covering a diverse range of domains. We independently generate attribute for each dataset.

	For range filtering queries, we evaluate 3 widely used real-world datasets: Deep~\cite{yandex_deep_dataset}, YT-Audio~\cite{youtube8m_dataset}, and WIT~\cite{wit_dataset}. Following prior work~\cite{DSG}, we use the data point ID within each dataset as its attribute. This ensures that the data is initially sorted.

@@

	As shown in Figure~\ref{fig:exp_4_1_MultiLabel_1thread}, UNG achieves the best performance across all datasets. For IVF-based methods not specifically optimized for hybrid queries (Faiss\_IVF and Milvus), their QPS remains relatively stable as recall varies. Milvus, \textcolor{red}{as a general-purpose system}, exhibits comparatively poor performance. In contrast, Faiss—especially when combined with HQI\_Batch—achieves medium-level performance. Its original IVF implementation naturally supports pre-filtering and does not suffer from the complexity introduced by an increased number of attributes. In fact, filtering can reduce the effective candidate set, improving efficiency.
%	As shown in Figure~\ref{fig:exp_4_1_MultiLabel_1thread}, UNG achieves the best performance across all datasets. For IVF-based methods that are not specifically optimized for hybrid query scenarios (Faiss\_IVF and Milvus), their QPS remains relatively stable as the recall varies. Milvus, being a general-purpose system, exhibits comparatively poor performance. In contrast, Faiss—especially when combined with HQI\_Batch—achieves medium-level performance. Its original IVF implementation naturally supports pre-filtering and does not suffer from the complexity introduced by an increased number of attributes. In fact, filtering can reduce the effective candidate set, improving efficiency.

	The IVF-based algorithm CAPS, optimized for hybrid queries, employs a multi-level partitioning strategy that introduces additional overhead. This overhead negatively impacts performan-\\ce on small datasets. However, on large datasets, the overhead is amortized, and the benefits of attribute-aware partitioning become evident, making CAPS the second-best performer after UNG.

%	NHQ exhibits mediocre performance overall and performs particularly poorly on small datasets. It relies on fusion distance calculations. The fusion distance computation is highly sensitive to the number of attributes involved, thus impacting both efficiency and accuracy under multi-attribute filtering.
	NHQ \textcolor{red}{performs moderately overall and poorly on small datasets}. Its reliance on fusion distance computations \textcolor{red}{makes it highly sensitive to the number of attributes}, impacting both efficiency and accuracy under multi-attribute filtering.

@@

	\textit{\textbf{Different Datasets.}}  
	To evaluate the performance of various methods across different datasets, we conduct experiments on 7 datasets, including one dataset for out-of-distribution (OOD) queries. Our analysis focuses on performance variations concerning dataset size, LID, and vector dimensionality.

@@

	\begin{figure*}[h]
		\centering
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_8_2.pdf}
		\caption{RF-ANN search }
		\label{fig:exp_8_2}
	\end{figure*}

@@

	% \begin{figure*}[htbp]
		%     \centering
		%     \setlength{\abovecaptionskip}{0cm}
		%     \setlength{\belowcaptionskip}{-0.4cm}
		%     \includesvg[width=\textwidth]{figures/exp/exp_8_3.svg}
		%     \caption{Dataset Effect on RF-ANN Search}
		%     \label{fig:exp_8_3}
		% \end{figure*}
	\begin{figure*}[htbp]
		\centering
		    \setlength{\abovecaptionskip}{0cm}
		    \setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.75\textwidth]{figures/exp/exp_8_3.pdf}
		\caption{Dataset Effect on RF-ANN Search}
		\label{fig:exp_8_3}
	\end{figure*}

>>>>

	\noindent\textbf{\underline{UNG \cite{UNG}.}} The above simultaneous filtering algorithms may not guarantee the completeness of the results and perform poorly when the attribute selectivity is low. UNG is proposed to overcome these limitations.
	UNG is a unified framework that integrates diverse graph-based ANN indexes for hybrid query. It first groups the dataset by attribute sets, ensuring that vectors in each group share identical attributes. Then, it constructs a Label Navigating Graph (LNG) to encode inclusion relationships among attribute sets. Within each group, UNG builds graph-based ANN indexes (e.g., Vamana, HNSW) and connects them via cross-group edges to enable efficient cross-group search.
	%UNG supports Boolean filtering with 2 modes:  1) The query attribute set is a subset of the data attribute set.  2) The query attribute set exactly matches the data attribute set.

@@

	For attribute filtering tasks, we employ 6 real-world datasets widely adopted in existing literature: Msong~\cite{msong2011}, Audio~\cite{audio_unknown}, SIFT1M~\cite{sift2010}, GIST1M~\cite{sift2010}, GloVe~\cite{GloVe2015}, Enron~\cite{enron2015} , covering a diverse range of domains. We independently generate attribute for each dataset.
	For range filtering queries, we evaluate 3 widely used real-world datasets: Deep~\cite{yandex_deep_dataset}, YT-Audio~\cite{youtube8m_dataset}, and WIT~\cite{wit_dataset}. Following prior work~\cite{DSG}, we use the data point ID within each dataset as its attribute. This ensures that the data is initially sorted.
	\textcolor{red}{In addition, we evaluate the performance of different algorithms on the Out-of-Distribution (OOD) dataset Text2Image~\cite{texttoimage}.}

@@

@@

	As shown in Figure~\ref{fig:exp_4_1_MultiLabel_1thread}, UNG achieves the best performance across all datasets. For IVF-based methods that are not specifically optimized for hybrid query scenarios (Faiss\_IVF and Milvus), their QPS remains relatively stable as the recall varies. Milvus, being a general-purpose system, exhibits comparatively poor performance. In contrast, Faiss—especially when combined with HQI\_Batch—achieves medium-level performance. Its original IVF implementation naturally supports pre-filtering and does not suffer from the complexity introduced by an increased number of attributes. In fact, filtering can reduce the effective candidate set, improving efficiency.

	The IVF-based algorithm CAPS, optimized for hybrid query scenarios, employs a multi-level partitioning strategy that introduces additional overhead. This overhead negatively impacts performan-\\ce on small datasets. However, on large datasets, the overhead is amortized, and the benefits of attribute-aware partitioning become evident, making CAPS the second-best performer after UNG.

	NHQ exhibits mediocre performance overall and performs particularly poorly on small datasets. It relies on fusion distance calculations. The fusion distance computation is highly sensitive to the number of attributes involved, thus impacting both efficiency and accuracy under multi-attribute filtering.

@@

	\textit{\textbf{Different Datasets.}}  
	To evaluate the performance of various methods across different datasets, we conduct experiments on 7 datasets, including one dataset Text2Image for OOD queries. Our analysis focuses on performance variations concerning dataset size, LID, and vector dimensionality.

@@

	\begin{figure*}[htbp]
		\centering
		  \setlength{\abovecaptionskip}{0cm}
		    \setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_8_2.pdf}
		\caption{RF-ANN search }
		\label{fig:exp_8_2}
	\end{figure*}

@@

	% \begin{figure*}[htbp]
		%     \centering
		%     \setlength{\abovecaptionskip}{0cm}
		%     \setlength{\belowcaptionskip}{-0.4cm}
		%     \includesvg[width=\textwidth]{figures/exp/exp_8_3.svg}
		%     \caption{Dataset Effect on RF-ANN Search}
		%     \label{fig:exp_8_3}
		% \end{figure*}
	\begin{figure*}[htbp]
		\centering
		    \setlength{\abovecaptionskip}{0cm}
		    \setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.8\textwidth]{figures/exp/exp_8_3.pdf}
		\caption{Dataset Effect on RF-ANN Search}
		\label{fig:exp_8_3}
	\end{figure*}
